---
title: "EPLC data-preprocessing"
output: 
    html_notebook:
      toc: yes
      theme: simplex
      toc_float: true
      
---

```{r setup, include = F}
library(data.table)
library(ggplot2)
library(lubridate)
library(lme4)
library(rethinking)
library(magrittr)
library(stringr)

knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
```

I pre-process here EPLC data from 2014 and 2015. 

## Data Summaries and Structure

```{r data preprocessing}
eplc2014 <- fread("eplcproj2014.csv", encoding = "UTF-8")
eplc2015 <- fread("eplcproj2015.csv", encoding = "UTF-8")
eplcbind <- unique(rbindlist(list(eplc2014, eplc2015)))

## transforming the location variables to lower case

head(eplcbind, 100)

```

```{r}
summary(eplcbind)
str(eplcbind)
```

Add column for total project duration, which is derived from subtracting the actual completion dates to the actual start dates:

```{r calculate actual start and completion dates}
month_conv = function(int) {
  ifelse(int/10 < 1, paste(0, int, sep = ''), as.character(int))
}

eplcbind[,actual_start_date := paste(actual_start_year, month_conv(actual_start_month), 
                                     '01',sep = '')]
eplcbind[,actual_start_date:=ymd(actual_start_date)]
eplcbind[,actual_completion_date := paste(actual_completion_year, month_conv(actual_completion_month), 
                                     '01',sep = '')]
eplcbind[,actual_completion_date:=ymd(actual_completion_date)]
eplcbind[,project_duration := floor(as.integer(actual_completion_date - actual_start_date)/30)]
eplcbind[,project_duration_dy := as.integer(actual_completion_date - actual_start_date)]
head(eplcbind[,.(actual_start_date, actual_completion_date, project_duration)])
```

Project duration is in months.
```{r}
summary(eplcbind[,.(project_duration)])
```

I also converted the contract start time to date format, and derived a variable `lag_time` which is the duration between the start of contract and the actual start time.

```{r}
eplcbind[,contract_start_date_ymd := as_date(mdy_hm(contract_start_date))]
eplcbind[,lag_time:=actual_start_date - contract_start_date_ymd]
head(eplcbind[,.(contract_start_date_ymd, actual_start_date, contract_start_date_ymd, lag_time)])

```

Shown below, lag time are negative values which means that this could be just related to the actual start date.

```{r}
summary(eplcbind[,.(as.integer(lag_time))])
```

## Deduplicating rows

There is significant duplication of entries in the row binded data frames for eplc 2014 to 2015. 

```{r}
eplcbind[,cnt:=.N, project_id][cnt > 1][order(project_id)]
```

A simple omission of missing values made the rows distinct from each other according to the project_ids:

```{r}
eplc_dedup <- na.omit(eplcbind)

nrow(eplc_dedup) == eplc_dedup[,uniqueN(project_id)]
```

## Project cost exploration

```{r}
ggplot(eplc_dedup, aes(project_cost)) +
  geom_histogram(bins = 30)  +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_bw()
```

```{r}
ggplot(eplc_dedup, aes(log(project_cost + 1))) +
         geom_histogram(bins = 50) +
  scale_x_continuous(limits = c(2.5,15)) +
  theme_bw()
```


Shown below is a histogram of the project cost with increased granularity (bins = 1000). A lot of peaks can be seen with increasing project cost.

```{r}
ggplot(eplc_dedup, aes(project_cost)) +
  geom_histogram(bins = 500) +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_bw()
```

Not much to be said about the plot below, when we use the raw project cost.

```{r}
## Relating project cost and total project duration
ggplot(eplc_dedup, aes(x = project_duration, y = project_cost))+
 geom_point(alpha = 0.1) +
 scale_x_continuous(limits = c(0,25)) +
  scale_y_continuous(breaks = seq(0, 5e6, by = 3e5)) +
  theme_bw()
```


taking the log of the project cost gives a more discernible pattern compared to just using the raw value as shown below.
```{r, warning=F, message = F}
ggplot(eplc_dedup[!is.na(project_duration)], aes(y = log(project_cost + 10), x = project_duration)) + 
  # geom_point( color = "dodgerblue4", alpha = 0.05 , shape = 1, position = "jitter", width = 0.05, 
  #             height = 0.05) + 
  geom_jitter(width = .5, height = .1 , color = "dodgerblue4", alpha = 0.05) +
  geom_smooth(formula = y ~ x, color = "tomato1", se  =F) +
  theme_minimal() +
  scale_x_continuous(limits = c(0,24), breaks = seq(0, 24, by = 2)) +
  scale_y_continuous(limits = c(0, 15), breaks = seq(0,15, by = 3)) + 
 labs(x = "project duration in months", y = "log project cost") +
  ggtitle("log project cost vs project duration")
```

## Exploration on Funding source type

There are a total of 153 types of funding sources.

```{r}
eplc_dedup[,.N, by=fs_type] %>%
  ggplot(aes(N)) +
  geom_histogram(bins = 100, fill = "dodgerblue4", alpha =0.6) + 
  theme_minimal() 
```



Plotting the median costs by funding source, it can be observed that there is considerable variation across funding sources. 

```{r}
eplc_dedup[,.(median_cost = median(project_cost, na.rm = T), 
            count = .N), by=fs_type] %>% 
  .[order(-median_cost)] %>%
  ggplot(aes(x= median_cost)) +
  geom_histogram(bins = 50 , fill = "dodgerblue4", alpha = 0.6) +
  theme_minimal() + 
  ggtitle("Median costs histogram of funding source types") +
  labs(x = 'median cost')
```

```{r}
eplc_dedup[,log_project_cost := log(project_cost + 1)]
data_mm <- na.omit(eplc_dedup[,.(log_project_cost, fs_type, project_id)])
mm_fs_type <- lmer(log_project_cost ~ (1|fs_type), data = data_mm)

data_mm[,log_project_cost_est := predict(mm_fs_type)]
data_mm[,project_cost_est := exp(log_project_cost_est) + 1] 

fstypeCostEstimation <- 
unique(data_mm[,.(project_cost_est, fs_type)]) %>%
  .[order(-fs_type)] %>%
  .[
    eplc_dedup[,.(
        project_cost_mean_emp = mean(project_cost, na.rm = T),
        count = .N
        ), 
      by = fs_type
      ],
    nomatch = 0, 
    on = 'fs_type'] %>%
  .[order(-project_cost_est)]

fstypeCostEstimation

```

```{r}
ggplot(fstypeCostEstimation, aes(x=project_cost_est)) +
  geom_histogram(bins = 50) 
```





## Exploration on location columns

We explore here the columns `project_description`, `project_location` and `implementing_office` 

```{r}
eplc_dedup[,lapply(.SD, uniqueN), .SDcols = c("project_location", "implementing_office", "project_description")]
```

Let's start first with the smallest number of distinct values `implement_office`
```{r}
eplc_dedup[,.N, implementing_office]
```


```{r}
eplcbind[,.N, project_location]
```



## Exploration by region population

We need to join first the corresponding population data for the different regions. This population data was scraped from psa.gov.ph. 

```{r}
PopnPerGovUnit <- readRDS("Population_per_gov_unit.RDS") 

#Transform string variables into lower case
StringVars <- c("Region", "Province", "City_Municipality", "location")
PopnPerGovUnit[,c(StringVars):=lapply(.SD, str_to_lower),.SDcols = eval(StringVars)]
PopnPerGovUnit[,locOriginal:= location]

# Remove words inside parenthesis
PopnPerGovUnit[,location := str_replace_all(location, "\\(.*\\)", "")]

# remove numbers in City and municipal names
PopnPerGovUnit[level == "City_Municipality", location := str_replace_all(location, "\\d|\\d\\s$", "")]

# remove punctuation marks
PopnPerGovUnit[,location := str_replace_all(location, "[[:punct:]]", "")]
PopnPerGovUnit[,location := str_replace_all(location, "^\\s|\\s+$", "") %>% 
                 #remove more than one white space
                 str_replace_all("\\s{2,}", " ")]

### EDGE CASES ####
# changing some instances of "city of" into just "<name> city, retain the island garden city of Samal and Science city of Muñoz as these are legitimate names"
PopnPerGovUnit[!str_detect(location,"island garden|science city"), 
               location:= str_replace_all(location, "city of (.*)$", "\\1 city")] 
# Change caraga to region XIII
PopnPerGovUnit[location == "caraga" & is.na(Region), location := "region xiii"]
# interchange occidental and oriental in mindoro i.e. instead of occidental mindoro, mindor occidental 
PopnPerGovUnit[str_detect(location, "[occidental, oriental] mindoro"),
               location := str_replace_all(location, "(.+) mindoro", "mindoro \\1")]

# Remove provinces in the national capital region 
PopnPerGovUnit[Region == "national capital region", Province := NA]

# add a District column to accomodate the lack of Provinces in the national capital region
PopnPerGovUnit[City_Municipality %in% c("tondo", "binondo", "santa cruz", "san nicolas",
                                        "ermita", "malate", "pandacan", "santa ana",
                                        "san miguel", "intramuros", "paco", "port area",
                                        "quiapo") & Region =="national capital region",
               City_Municipality := "manila"] 
PopnPerGovUnit[City_Municipality == "manila"]



# A special is case is Manila, where it is subdivided into corresponding legislative districts. But in this data frame, Manila is subdivided into its geographic districts (Tondo, Binondo, Sta. Cruz etc.) 
# district 1 brgy 1-146
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 1:146) &
                 level == "barangay", District := "manila district 1"]
# district 2 147-267
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 147:267) &
                 level == "barangay", District := "manila district 2"]
# district 3 268-394
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 268:394) &
                 level == "barangay", District := "manila district 3"]
# district 4 395-586
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 395:586) &
                 level == "barangay", District := "manila district 4"]
# district 5 649-828
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 649:828) &
                 level == "barangay", District := "manila district 5"]
# district 6 587-648, 829-905
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% c(587:648, 829:905)) &
                 level == "barangay", District := "manila district 6"]


```


```{r}
unique(PopnPerGovUnit[,.(location)])
```


For the uppermost level of locations, which is not necessarily a Region/Province, I removed the string "district engineering office" and also removed the ordinal numbers. Other location vars such as project_description and project_location are transformed to lower cases.

```{r}
eplc_dedup <- na.omit(eplcbind)
locVars <- c("implementing_office", "project_location", "project_description")
eplc_dedup[,c(locVars) := lapply(.SD, str_to_lower), .SDcols = eval(locVars)]
eplc_dedup[,implementing_office:=str_to_lower(implementing_office)]
eplc_dedup[,UpperLevel:=str_replace_all(implementing_office, "district engineering office", "") %>%
           str_replace_all("\\s+\\d{1}\\w+", "") %>%
           str_replace_all("\\s+$", "")]

#### EDGE CASES #####
# fix ñ characters
eplc_dedup[str_detect(project_location , "\\\x96"),
           project_location := str_replace_all(project_location, "\\\x96", "ñ")]
# changing some location strings to more relevant ones
eplc_dedup[UpperLevel == "mt.province", UpperLevel := "mountain province"]

# change dinagat island to dinagat islands
eplc_dedup[UpperLevel == "dinagat island", UpperLevel := "dinagat islands"]

# remove punctuations in region iv upper level
eplc_dedup[str_detect(UpperLevel, "region iv"),
           UpperLevel := str_replace_all(UpperLevel,"-", "")]

# change southern mindoro to oriental mindoro
eplc_dedup[UpperLevel == "southern mindoro", UpperLevel := "mindoro oriental"]

# remove punctations in project_location
eplc_dedup[,project_location := str_replace_all(project_location,"\\(.*\\)", " ") %>%
             str_replace_all("^\\s+|\\s+$", "") %>%
             str_replace("-", " ") %>%
             str_replace("[[:punct:]]", "") %>% 
             str_replace("\\s{2,}", "\\s")]


```


```{r}
unique(eplc_dedup[,.(project_location)]) 
  
  
  
```


At this point, I devise an algorithm to sequentially pinpoint the location, and attaching finally the population for a certain project_location:

The method goes like this:
+ search the corresponding region/city/province in the `implementing_office` column
+ From the found region/city/province, narrow the search and find the corresponding city if a province was found,
province if a region was found, and barangay if a city was found. This was then matched with the `project_location` column

```{r}
FindLoc = function(UpperLevel, project_location, project_description, pop_df) {
 
  project_location <- eplc_dedup$project_location
  pop_df <- PopnPerGovUnit
  project_description <- eplc_dedup$project_description
  UpperLevel <- eplc_dedup$UpperLevel 
  
  # determine the location from project_location
  projectLocUnique <- unique(project_location)
  loc <- pop_df$location
  FromProjLoc <-
    lapply(projectLocUnique, function(x)
    str_which(loc, paste0("\\b", x, "\\b")))
  FromProjLocDf <- data.table(project_location = projectLocUnique, 
                              FromProjLoc = FromProjLoc)
  df <- data.table(UpperLevel = UpperLevel, project_location = project_location, 
                   project_description = project_description)
  df[FromProjLocDf, Cnd := i.FromProjLoc, on = "project_location" ]
  
  Cnd <- df[42237, 4]
  UpperLevel <- df[42237, ][["UpperLevel"]]
 
 DetectBestLoc = function(Cnd, UpperLevel, pop_df)  {
  # Using candidate indices, find the most appropriate location by matching with the relevant Province, Region or 
   # City/Municipality in the implementing office section
   
    Cnd <- unlist(Cnd) 
    
    if (length(Cnd) == 0) {
      # if length of candidates is 0, go directly to searching UpperLevel
      
      RelevantData <- pop_df[str_detect(location, paste0("\\b", UpperLevel, "\\b")) &
                               level %in% c("City_Municipality", "Province", "Region"),
                               .(Province, Region, City_Municipality, location, population)] 
     
     # if regular expressions give multiple matches, try doing an exact match
     if (nrow(RelevantData) > 1) {
       RelevantData <- RelevantData[location == UpperLevel]
     }
     
     Loc <- RelevantData[,-5]
     BestMatchIndex <- 1
        
    }
    
   if (length(Cnd) > 0)  {
     RelevantData <-
       pop_df[Cnd, ] %>%
       .[, .(Region, Province, City_Municipality, location, population)]
      
      if (length(Cnd) == 1) {
       # if Cnd is equal to 1, it is most probable that it has only one match in the population df, and it is assumed
       # that it is reasonable to directly query its corresponding tuple (Province, Region, City_Municipality, Barangay)
        Loc <- RelevantData[,1:4]
        
      } else { 
      
      BestMatchIndex <- 
      lapply(RelevantData[,.(Province, Region, City_Municipality)], 
             function(x) str_detect(x, UpperLevel)) %>%
        as.data.table() %>%
        .[,sum_bool := apply(.SD, 1, sum, na.rm = T)] %>%
        .[["sum_bool"]] %>%
        which.max()
    
       Loc <- RelevantData[BestMatchIndex,1:4]
       
      }
   }
    
    # for edge cases where two locations are indistinguisable such as 
    #siquijor town and siquijor province, prefer province level hence:
    if (nrow(Loc) > 1) {
      Loc <- Loc[is.na(Province)]
    }
    
    #identify null columns for restructuring the necessary information
    NACols <- which(unname(sapply(Loc, is.na, USE.NAMES =  F))) 
    
     # removing null columns and relabeling the columns by hierarchy
    if (length(NACols) != 0) {
      Loc[, (NACols) := NULL]
      Loc[, (paste0("V", 1:length(NACols))) := NA]
      setnames(Loc, colnames(Loc),
               c("Region", "Province", 
                 "City_Municipality", "Barangay"))
    }
    
     # Attaching the population col
     return(cbind(Loc, RelevantData[BestMatchIndex, 5]))
   
 }
 
 for (i in 38634:nrow(df)) {
   df[i, BestLocation := list(mapply(
     function(x, y) DetectBestLoc(x, y , pop_df),
     Cnd,
     UpperLevel,
     SIMPLIFY = F
     ))]
   
 }
 
 
 
 df[i,]
 
 
 DetectBestLoc(df[1,4],df[1,1], pop_df)
 return(df$BestLocation)
  
  
}

eplc_dedup[1, BestLoc := FindLoc(UpperLevel, project_location, project_description, PopnPerGovUnit)]

eplc_dedup[str_detect(project_location,"city of")]

d <-
eplc_dedup[, .(project_id, project_location, project_description, location = UpperLevel)] %>%
  .[
  PopnPerGovUnit[level %in% c("City_Municipality", "Province"),.(location, level)],
  on = "location",
  allow.cartesian = T
  
]


str_detect("city of cebu capital", "(?=.*\\bcebu\\b)(?=.*?\\bcity\\b).*") 
  

```


```{r}
d[,cnt :=.N, project_id]
d[cnt > 1][order(project_id)] %>%
  .[,.(loc = unique(location))]
```


There seems to be some duplication with the places number of rows equal to `r nrow(PopnPerGovUnit)` while number of unique values equal to `r PopnPerGovUnit[,uniqueN(location)]`

```{r}
PopnPerGovUnit[,cnt:=.N, character] %>%
  .[cnt >1]
```


```{r}
PopnPerGovUnit[character == 'ilian']
```


Most of the duplications come from the brgy level

```{r}
PopnPerGovUnit[cnt > 1, .N, level]
```



Let us try doing a naive join between the location names in `eplcbind` and in `PopnPerGovUnit`:

```{r}
PopnPerGovUnit[,character:=stringi::stri_trans_tolower(character)]
```

```{r}
eplcbind[,project_location := stringi::stri_trans_tolower(project_location)]

setkey(eplcbind, project_location)
setkey(PopnPerGovUnit, character)
PopnPerGovUnit[eplcbind, ]

Eplcbind_w_Pop <- merge(PopnPerGovUnit, eplcbind, by.x = "character", 
                        by.y ="project_location", all.x=T)
Eplcbind_w_Pop
```




```{r, eval = F}
data_gamma <- na.omit(eplcbind[project_cost >0,.(project_cost, fs_type)])
data_gamma[,fs_type := as.factor(fs_type)]
data_gamma[,log_project_cost := log(project_cost)]
dt <- as.data.frame(data_gamma)
mdl <- map2stan(
  alist(
       log(mu) <- a_fs_type[fs_type],
      project_cost ~ dgamma2(mu, scale),
      # rate <- shape/exp(log_mu),
      a_fs_type[fs_type] ~ dnorm(a, sigma),
      a ~ dnorm(0,1),
      sigma ~ dcauchy(0,1),
      scale ~ dexp(2)
  ),
  data = dt,
  start = list(sigma = 3),
  constraints=list(sigma="lower=0", scale = "lower=1")
 
)
```
