---
title: "Data Structure and pre-processing"
output: 
    html_document:
      toc: yes
      theme: simplex
      toc_float: true
      
---

```{r setup, include = F}
library(data.table)
library(ggplot2)
library(lubridate)
library(lme4)
library(rethinking)
library(magrittr)
library(stringr)

knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
```

I pre-process here EPLC data from 2014 and 2015. 

# Data Summaries and Structure

```{r data preprocessing}
eplc2014 <- fread("eplcproj2014.csv", encoding = "UTF-8")
eplc2015 <- fread("eplcproj2015.csv", encoding = "UTF-8")
eplcbind <- unique(rbindlist(list(eplc2014, eplc2015)))

head(eplcbind)

```


Add column for total project duration, which is derived from subtracting the actual completion dates to the actual start dates:

```{r calculate actual start and completion dates}
month_conv = function(int) {
  ifelse(int/10 < 1, paste(0, int, sep = ''), as.character(int))
}

eplcbind[,actual_start_date := paste(actual_start_year, month_conv(actual_start_month), 
                                     '01',sep = '')]
eplcbind[,actual_start_date:=ymd(actual_start_date)]
eplcbind[,actual_completion_date := paste(actual_completion_year, month_conv(actual_completion_month), 
                                     '01',sep = '')]
eplcbind[,actual_completion_date:=ymd(actual_completion_date)]
eplcbind[,project_duration := floor(as.integer(actual_completion_date - actual_start_date)/30)]
eplcbind[,project_duration_dy := as.integer(actual_completion_date - actual_start_date)]
head(eplcbind[,.(actual_start_date, actual_completion_date, project_duration)])
```

Project duration is in months.

I also converted the contract start time to date format, and derived a variable `lag_time` which is the duration between the start of contract and the actual start time.

```{r}
eplcbind[,contract_start_date_ymd := as_date(mdy_hm(contract_start_date))]
eplcbind[,lag_time:=actual_start_date - contract_start_date_ymd]
head(eplcbind[,.(contract_start_date_ymd, actual_start_date, contract_start_date_ymd, lag_time)])

```

Shown below, lag time are negative values which means that this could be just related to the actual start date.

```{r}
summary(eplcbind[,.(as.integer(lag_time))])
```

## Deduplicating rows

There is significant duplication of entries in the row binded data frames for eplc 2014 to 2015. 

```{r, echo = T}
eplcbind[,cnt:=.N, project_id][cnt > 1][order(project_id)] %>%
  nrow()
```

A simple omission of missing values made the rows distinct from each other according to the project_ids:

```{r, echo = T}
eplc_dedup <- na.omit(eplcbind)
nrow(eplc_dedup) == eplc_dedup[,uniqueN(project_id)]
```

```{r, include = F}
saveRDS(eplc_dedup, "eplc_preproc.rds")
```




## Exploration by region population

We need to join first the corresponding population data for the different regions. This population data was scraped from psa.gov.ph. 




```{r}
d[,cnt :=.N, project_id]
d[cnt > 1][order(project_id)] %>%
  .[,.(loc = unique(location))]
```


There seems to be some duplication with the places number of rows equal to `r nrow(PopnPerGovUnit)` while number of unique values equal to `r PopnPerGovUnit[,uniqueN(location)]`

```{r}
PopnPerGovUnit[,cnt:=.N, character] %>%
  .[cnt >1]
```


```{r}
PopnPerGovUnit[character == 'ilian']
```


Most of the duplications come from the brgy level

```{r}
PopnPerGovUnit[cnt > 1, .N, level]
```



Let us try doing a naive join between the location names in `eplcbind` and in `PopnPerGovUnit`:

```{r}
PopnPerGovUnit[,character:=stringi::stri_trans_tolower(character)]
```

```{r}
eplcbind[,project_location := stringi::stri_trans_tolower(project_location)]

setkey(eplcbind, project_location)
setkey(PopnPerGovUnit, character)
PopnPerGovUnit[eplcbind, ]

Eplcbind_w_Pop <- merge(PopnPerGovUnit, eplcbind, by.x = "character", 
                        by.y ="project_location", all.x=T)
Eplcbind_w_Pop
```


