---
title: "Project Cost Exploration on location population"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(magrittr)
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.width = 6, fig.height = 4)
knitr::opts_knit$set(root.dir = "../")
theme_set(theme_minimal())
```

```{r include = F}
eplc_dedup <- readRDS("bin/eplc_preproc.rds")
PopnPerGovUnit_proc <- readRDS("bin/PopnPerGovUnit-proc.rds")
# locwpop <- readRDS("bin/eplc-w-pop.rds")
```


## Associating government units to a project location

I started with the collection of data from psa.gov.ph via web scraping to obtain the necessary population information in all government units ranging from the barangay to the region. This was done in a separate script. After obtaining said information, some pre-processing steps were done to ensure that locations are uniform between the scraped and eplc data.

Code below was used to pre-process the table for population per government unit: 

```{r, code = readLines("../src/PopPerGovUnit_proc.R"), echo = T , eval = F}
```

While this code snippet was used to further process eplc data location fields: 
```{r, code = readLines("../src/eplc-loc-proc.R"), echo = T, eval = F}

```


For the uppermost level of locations, which is not necessarily a Region/Province, I removed the string "district engineering office" and also removed the ordinal numbers. Other location vars such as project_description and project_location were transformed to lower cases.

<br>

### determining population where a project is located

At this point, I devise an algorithm to sequentially pinpoint the location, and attaching finally the population for a certain project_location.


```{r, code = readLines("../src/find-loc.R"), echo = T , eval = F}
```
