---
title: "Project Cost Exploration on location population"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(magrittr)
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.width = 6, fig.height = 4)
theme_set(theme_minimal())
```

```{r include = F}
eplc_dedup <- readRDS("eplc_preproc.rds")
PopnPerGovUnit <- readRDS("Population_per_gov_unit.RDS") 
```


# Associating City/Municipality, Region and Barangay to a project location

I started with the collection of data from psa.gov.ph via web scraping to obtain the necessary population information in all government units ranging from the barangay to the region. This was done in a separate script. After obtaining said information, 

```{r, echo = T}

#Transform string variables into lower case
StringVars <- c("Region", "Province", "City_Municipality", "location")
PopnPerGovUnit[,c(StringVars):=lapply(.SD, str_to_lower),.SDcols = eval(StringVars)]
PopnPerGovUnit[,locOriginal:= location]

# Remove words inside parenthesis
PopnPerGovUnit[,location := str_replace_all(location, "\\(.*\\)", "")]

# remove numbers in City and municipal names
PopnPerGovUnit[level == "City_Municipality", location := str_replace_all(location, "\\d|\\d\\s$", "")]

# remove punctuation marks
PopnPerGovUnit[,location := str_replace_all(location, "[[:punct:]]", "")]
PopnPerGovUnit[,location := str_replace_all(location, "^\\s|\\s+$", "") %>% 
                 #remove more than one white space
                 str_replace_all("\\s{2,}", " ")]

### EDGE CASES ####
# changing some instances of "city of" into just "<name> city, retain the island garden city of Samal and Science city of Muñoz as these are legitimate names"
PopnPerGovUnit[!str_detect(location,"island garden|science city"), 
               location:= str_replace_all(location, "city of (.*)$", "\\1 city")] 
# Change caraga to region XIII
PopnPerGovUnit[location == "caraga" & is.na(Region), location := "region xiii"]
# interchange occidental and oriental in mindoro i.e. instead of occidental mindoro, mindor occidental 
PopnPerGovUnit[str_detect(location, "[occidental, oriental] mindoro"),
               location := str_replace_all(location, "(.+) mindoro", "mindoro \\1")]

# Remove provinces in the national capital region 
PopnPerGovUnit[Region == "national capital region", Province := NA]

# add a District column to accomodate the lack of Provinces in the national capital region
PopnPerGovUnit[City_Municipality %in% c("tondo", "binondo", "santa cruz", "san nicolas",
                                        "ermita", "malate", "pandacan", "santa ana",
                                        "san miguel", "intramuros", "paco", "port area",
                                        "quiapo") & Region =="national capital region",
               City_Municipality := "manila"] 



# A special is case is Manila, where it is subdivided into corresponding legislative districts. But in this data frame, Manila is subdivided into its geographic districts (Tondo, Binondo, Sta. Cruz etc.) 
# district 1 brgy 1-146
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 1:146) &
                 level == "barangay", District := "manila district 1"]
# district 2 147-267
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 147:267) &
                 level == "barangay", District := "manila district 2"]
# district 3 268-394
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 268:394) &
                 level == "barangay", District := "manila district 3"]
# district 4 395-586
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 395:586) &
                 level == "barangay", District := "manila district 4"]
# district 5 649-828
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% 649:828) &
                 level == "barangay", District := "manila district 5"]
# district 6 587-648, 829-905
PopnPerGovUnit[City_Municipality == "manila" & 
                 (as.integer(unlist(str_extract(location, "\\d+"))) %in% c(587:648, 829:905)) &
                 level == "barangay", District := "manila district 6"]


```




For the uppermost level of locations, which is not necessarily a Region/Province, I removed the string "district engineering office" and also removed the ordinal numbers. Other location vars such as project_description and project_location are transformed to lower cases.

```{r}
eplc_dedup <- na.omit(eplcbind)
locVars <- c("implementing_office", "project_location", "project_description")
eplc_dedup[,c(locVars) := lapply(.SD, str_to_lower), .SDcols = eval(locVars)]
eplc_dedup[,implementing_office:=str_to_lower(implementing_office)]
eplc_dedup[,UpperLevel:=str_replace_all(implementing_office, "district engineering office", "") %>%
           str_replace_all("\\s+\\d{1}\\w+", "") %>%
           str_replace_all("\\s+$", "")]

#### EDGE CASES #####
# fix ñ characters
eplc_dedup[str_detect(project_location , "\\\x96"),
           project_location := str_replace_all(project_location, "\\\x96", "ñ")]
# changing some location strings to more relevant ones
eplc_dedup[UpperLevel == "mt.province", UpperLevel := "mountain province"]

# change dinagat island to dinagat islands
eplc_dedup[UpperLevel == "dinagat island", UpperLevel := "dinagat islands"]

# remove punctuations in region iv upper level
eplc_dedup[str_detect(UpperLevel, "region iv"),
           UpperLevel := str_replace_all(UpperLevel,"-", "")]

# change southern mindoro to oriental mindoro
eplc_dedup[UpperLevel == "southern mindoro", UpperLevel := "mindoro oriental"]

# remove punctations in project_location
eplc_dedup[,project_location := str_replace_all(project_location,"\\(.*\\)", " ") %>%
             str_replace_all("^\\s+|\\s+$", "") %>%
             str_replace("-", " ") %>%
             str_replace("[[:punct:]]", "") %>% 
             str_replace("\\s{2,}", "\\s")]


```


```{r}
unique(eplc_dedup[,.(project_location)]) 
  
  
  
```


At this point, I devise an algorithm to sequentially pinpoint the location, and attaching finally the population for a certain project_location:

The method goes like this:
+ search the corresponding region/city/province in the `implementing_office` column
+ From the found region/city/province, narrow the search and find the corresponding city if a province was found,
province if a region was found, and barangay if a city was found. This was then matched with the `project_location` column

```{r}
FindLoc = function(UpperLevel, project_location, project_description, pop_df) {
 
  project_location <- eplc_dedup$project_location
  pop_df <- PopnPerGovUnit
  project_description <- eplc_dedup$project_description
  UpperLevel <- eplc_dedup$UpperLevel 
  
  # determine the location from project_location
  projectLocUnique <- unique(project_location)
  loc <- pop_df$location
  FromProjLoc <-
    lapply(projectLocUnique, function(x)
    str_which(loc, paste0("\\b", x, "\\b")))
  FromProjLocDf <- data.table(project_location = projectLocUnique, 
                              FromProjLoc = FromProjLoc)
  df <- data.table(UpperLevel = UpperLevel, project_location = project_location, 
                   project_description = project_description)
  df[FromProjLocDf, Cnd := i.FromProjLoc, on = "project_location" ]
  
  # Cnd <- df[17, 4]
  # UpperLevel <- df[17, ][["UpperLevel"]]
 
 DetectBestLoc = function(Cnd, UpperLevel, pop_df)  {
  # Using candidate indices, find the most appropriate location by matching with the relevant Province, Region or 
   # City/Municipality in the implementing office section
   
    Cnd <- unlist(Cnd) 
    
    if (length(Cnd) == 0) {
      # if length of candidates is 0, go directly to searching UpperLevel
      
      RelevantData <- pop_df[str_detect(location, paste0("\\b", UpperLevel, "\\b")) &
                               level %in% c("City_Municipality", "Province", "Region"),
                               .(Province, Region, City_Municipality, location, population)] 
     
     # if regular expressions give multiple matches, try doing an exact match
     if (nrow(RelevantData) > 1) {
       RelevantData <- RelevantData[location == UpperLevel]
     }
     
     Loc <- RelevantData[,-5]
     BestMatchIndex <- 1
        
    }
    
   if (length(Cnd) > 0)  {
     RelevantData <-
       pop_df[Cnd, ] %>%
       .[, .(Region, Province, City_Municipality, location, population)]
      
      if (length(Cnd) == 1) {
       # if Cnd is equal to 1, it is most probable that it has only one match in the population df, and it is assumed
       # that it is reasonable to directly query its corresponding tuple (Province, Region, City_Municipality, Barangay)
        Loc <- RelevantData[,1:4]
        
      } else { 
      
      BestMatchIndex <- 
      lapply(RelevantData[,.(Province, Region, City_Municipality)], 
             function(x) str_detect(x, UpperLevel)) %>%
        as.data.table() %>%
        .[,sum_bool := apply(.SD, 1, sum, na.rm = T)] %>%
        .[["sum_bool"]] %>%
        which.max()
    
       Loc <- RelevantData[BestMatchIndex,1:4]
       
      }
   }
    
    # for edge cases where two locations are indistinguisable such as 
    #siquijor town and siquijor province, prefer province level hence:
    if (nrow(Loc) > 1) {
      Loc <- Loc[is.na(Province)]
    }
    
    if (nrow(Loc) == 0) {
      return(data.table(Region = NA, Province = NA, City_Municipality = NA,
                                          Barangay = NA, population = NA))
    }
    #identify null columns for restructuring the necessary information
    NACols <- which(unname(sapply(Loc, is.na, USE.NAMES =  F))) 
    
     # removing null columns and relabeling the columns by hierarchy
    if (length(NACols) != 0) {
      Loc[, (NACols) := NULL]
      Loc[, (paste0("V", 1:length(NACols))) := NA]
      setnames(Loc, colnames(Loc),
               c("Region", "Province", 
                 "City_Municipality", "Barangay"))
    }
    
     # Attaching the population col
     return(cbind(Loc, RelevantData[BestMatchIndex, 5]))
   
 }
 
   df[, BestLocation := list(mapply(
     function(x, y) DetectBestLoc(x, y , pop_df),
     Cnd,
     UpperLevel,
     SIMPLIFY = F
     ))]
   bestLoc <- df$BestLocation
   
   df[,BestLocation := bestLoc]
   df[cbind(df[,.(project_id)], rbindlist(df$BestLocation)), 
      ':='(
        Region = i.Region,
        Province = i.Province,
        City_Municipality = i.City_Municpality,
        Barangay = i.Barangay,
        population = i.population
      ), on = "project_id"]
   df[,BestLocation := NULL]
   
 }
 
 
 
 d <- cbind(df, rbindlist(df$BestLocation))
 d
 
 DetectBestLoc(df[1,4],df[1,1], pop_df)
 return(df$BestLocation)
  
  
}

eplc_dedup[1, BestLoc := FindLoc(UpperLevel, project_location, project_description, PopnPerGovUnit)]

eplc_dedup[str_detect(project_location,"city of")]

d <-
eplc_dedup[, .(project_id, project_location, project_description, location = UpperLevel)] %>%
  .[
  PopnPerGovUnit[level %in% c("City_Municipality", "Province"),.(location, level)],
  on = "location",
  allow.cartesian = T
  
]


str_detect("city of cebu capital", "(?=.*\\bcebu\\b)(?=.*?\\bcity\\b).*") 
  

```
